---
title: "Data Processing"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r load_packages}
library(raster)
library(tidyverse)
```

## General details

We are creating a dataset of social-ecological covariates at a range of spatial resolutions: 1km, 5km, 10km, 25km, 50km, 100km. All data are open. 

## Land-cover data

We use [Corine Land Cover 2012](https://land.copernicus.eu/pan-european/corine-land-cover/clc-2012) at 100m resolution to calculate land-cover diversity at each resolution. 

## Elevation data

We use [EU-DEM v1.1](https://land.copernicus.eu/pan-european/satellite-derived-products/eu-dem/eu-dem-v1.1/view) at 30m resolution to calculate topographic variability at each resolution. 

## Population data

We use [Population density disaggregated with Corine land cover 2000](https://www.eea.europa.eu/data-and-maps/data/population-density-disaggregated-with-corine-land-cover-2000-2) at xx resolution to calculate total population at each resolution. 