---
title: "Flickr Models"
output: github_document
bibliography: C:/Users/lg1u16/DATA/BIBLIOGRAPHY/library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r load_packages}
library(hier.part)
library(MASS)
library(raster)
library(sf)
library(GGally)
library(broom)
library(MuMIn)
library(knitr)
library(tidyverse)
library(patchwork)

# resolutions
rlns <- c("1km", "5km", "10km", "25km", "50km", "100km")

# study extent (for now this is England, will need extending to Europe)
study_ext_sf <- st_read("~/DATA/ADMINISTRATIVE/gb_shapefile/GBR_adm1.shp", quiet = TRUE) %>% 
  filter(NAME_1 == "England") %>% 
  st_transform(crs = 3035)

study_ext <- as(study_ext_sf, "Spatial")

theme_set(theme_bw(base_size = 10) + theme(strip.background = element_blank(), 
                             panel.grid.major = element_blank(),
                             panel.grid.minor = element_blank()))
```

## Data collating

We're making one big dataframe of the response (Flickr photo locations) and covariates (proportion of non-built land covers, land-cover diversity, mean elevation, topographic variation and population density) at each of the analysis resolutions. 

```{r get_data}
get_df <- function(rln) {
  # list all files of specified resolution
  fnames <- list.files("data/covariates", pattern = paste0("_", rln), full.names = TRUE)
  fnames <- c(fnames, paste0("data/response/eng_flickr_", rln, ".tif"))
  
  # stack them
  dat <- stack(fnames)
  
  # crop by study extent
  dat <- crop(dat, study_ext)
  dat <- mask(dat, study_ext)
  
  # get into a dataframe
  df <- as.data.frame(dat, xy = TRUE) %>% 
    rename_all(funs(str_replace_all(., paste0("_", rln), ""))) %>% 
    na.omit %>% 
    mutate(resolution = rln)
  
  return(df)
}

# dataframe with obs for all data
df <- map_dfr(rlns, get_df) %>% 
  mutate(resolution = factor(resolution, 
                             levels = c("1km", 
                                        "5km", 
                                        "10km", 
                                        "25km", 
                                        "50km", 
                                        "100km")))
```

## Data exploration

First, let's just see whether there is sufficient variability and n in each resolution to fit models. 

```{r flickr_distribution}
ggplot(df, aes(x = eng_flickr)) + 
  geom_histogram() + 
  facet_wrap(~resolution, scales = "free")
```

```{r flickr_n}
df %>% 
  group_by(resolution) %>% 
  summarise(count = n())
```

Based on the distributions, we can fit models for all resolutions, but for the larger two (50km and 100km), n is very small. 

We can also inspect the spatial coverage of the Flickr data. 

```{r flickr_spatial}
ggplot(df, aes(x = x, y = y)) + 
  geom_raster() + 
  coord_equal() + 
  facet_wrap(~resolution) + 
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

Good coverage at all but 1km. Even that is better than the MENE data though and worth seeing if we can use it (n = 26715 rather than n = 5021). 

Finally, lets look at the relationships between variables:

```{r var_relationships}
p <- list()
for(r in rlns) {
  p[[r]] <- ggpairs(df %>% filter(resolution == r) %>% select(-resolution, -x, -y), progress = FALSE)
}

p
```

Seems that it's population and proportion of non-built land covers that are the strongest predictor at all resolutions, and that this effect increases with spatial resolution. LC diversity is next, with similar effect. Note that population density and proportion of non-built land covers are highly correlated at all resolutions except for 50km and 100km. 

## Statistical Models

We will fit negative binomial GLMs (due to overdispersion) with proportion of non-built land cover (clc_prop), mean elevation (dem_mean) and population (pop) and their first-order interactions as the full model. 

```{r flickr_mods}
fit_mod <- function(dat) {
  glm(eng_flickr ~ clc_agri + clc_forest + clc_shei + clc_veg + clc_water + clc_wetland + pop,
      family = "poisson",
      data = dat)
}

part_var <- function(dat) {
  resp <- dat %>% pull(eng_flickr)
  pred <- dat %>% dplyr::select(clc_agri, clc_forest, clc_shei, clc_water, clc_wetland, pop)
  
  part <- hier.part(resp, pred, family = "poisson", gof = "logLik", barplot = FALSE)$I.perc
  
  out <- tibble(var = rownames(part), expl = part$I)
}

scale_cols <- function(x) {
  scale_this <- function(y) as.vector(scale(y))
  #x <- x %>% mutate(pop = log(pop + 1))
  mutate_at(x, .vars = vars(-eng_flickr), .funs = funs(scale_this))
}

# get the CA function
source("../upscaling/R/ca_glm.R")

# get all of the required stats
get_mod_stats <- function(mod) {
  coeff <- coef(mod)[-1] %>% 
    data.frame %>% 
    as_tibble(rownames = "variable") %>% 
    rename(coef = 2)
  
  conf <- confint(mod)[-1,] %>% 
    data.frame %>% 
    as_tibble(rownames = "variable") %>% 
    rename(lci = X2.5.., uci = X97.5..)
  
  ca <- calc_commonality(mod)$CCTotalbyVar %>% 
    as_tibble(rownames = "variable")
  
  out <- inner_join(coeff, conf) %>% 
    inner_join(ca) %>% 
    mutate(
      r_s = map_dbl(variable, function(x) {
        out <- cor(model.matrix(mod)[,x], mod$fitted.values)
      }),
      r_s2 = r_s^2)
  
  return(out)
}

mod <- df %>% 
  group_by(resolution) %>% 
  nest() %>% 
  mutate(data_scaled = map(data, scale_cols),
         mod = map(data_scaled, fit_mod),
         mod_glance = map(mod, glance),
         mod_stats = map(mod, get_mod_stats),
         mod_part = map(data_scaled, part_var),
         mod_resid = map(mod, resid),
         mod_fitted = map(mod, fitted))

mod_D2 <- mod %>%
  dplyr::select(resolution, mod_glance) %>% 
  unnest() %>% 
  mutate(D2 = (null.deviance - deviance)/null.deviance) %>% 
  dplyr::select(resolution, D2)

mod_part <- mod %>% 
  dplyr::select(resolution, mod_part) %>% 
  unnest()

mod_stats <- mod %>%
  dplyr::select(resolution, mod_stats) %>% 
  unnest()
```

TODO: Testing residuals with DHARMa showed overdispersion with strong patterns in the residuals for CLC and POP. Would be worth fitting a quadratic term for each of these. Even with population logged, there are some low population cells (even at 100km level) with high residual values. 

How well do they fit? NB We are using $D^2$, which is equivalent to $R^2$ [@Guisan2000a]. Note also that due to the differences in sample size, $D^2$ values are not directly comparable.

```{r}
ggplot(mod_D2, aes(x = resolution, y = D2, group = 1)) + 
  geom_point() + 
  stat_summary(fun.y=sum, geom="line")
```

Explanatory power of the model increases greatly between 1km and 5km resolutions (with increasing explanatory power as resolution increases, as to be expected). Might be best to drop the 1km model. 

What about the relative importances?

NB This plot shows the proportion of variance explained by each variable of the total variance explained. It's greater than 1 because there are shared explained variances

```{r}
p1 <- ggplot(mod_D2, aes(x = resolution, y = D2)) + 
  geom_bar(stat = "identity") + 
  labs(x = "", y = expression(D^{2})) + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

p2 <- ggplot(mod_part, aes(x = resolution, y = expl, fill = var)) + 
  scale_fill_viridis_d(name = "Variable") + 
  geom_bar(stat = "identity", position = "stack") + 
  labs(x = "Resolution", y = "Variance explained %")

p1 + p2 + 
  plot_layout(ncol = 1, heights = c(1, 3)) + 
  plot_annotation(tag_levels = "a", tag_suffix = ")")
```

- Scale dependencies seem a bit more clear cut with this one (than MENE data).
- Importance of agriculture decreases with resolution
- Importance of population increases with resolution
- there's a quadratic effect of diversity (important at mid-level resolutions - is this where it is captured? 1km too small to capture, 100km there is far more within- than among-unit variation?)
- Water important at 1km, but then not until 50km.
- Forest really not important at all but the coarsest resolutions. 
- $D^2$ increase more stable/linear than with the MENE data

What about the relationships?

```{r}
ggplot(mod_stats, aes(x = resolution, y = coef)) + 
  geom_point() + 
  geom_hline(yintercept = 0) + 
  geom_errorbar(aes(ymin = lci, ymax = uci)) + 
  facet_wrap(~variable, nrow = 1)
```

- Agriculture always negative
- Forest often negative except at 100km
- Quadratic effect with diversity (strongest relationship at mid-resolutions)
- Vegetation always negative
- Water becomes negative at 10km and then becomes very negative at highest resolutions (likely an effect of losing high % water cells?)
- Population always positive, but the strength varies. 

## Bright & dark spots

Following Frei et al. [-@Frei2018] and Cinner et al. [-@Cinner2016], we calculate bright spots and dark spots. We define bright/dark spots as those where the observed value is +/- 1SD (of observed values) from the expected value. 

```{r, fig.width = 15, fig.height = 10}
bd_spots <- mod %>% 
  select(resolution, data, mod_fitted, mod_resid) %>% 
  unnest() %>% 
  group_by(resolution) %>% 
  mutate(diff = mod_fitted - eng_flickr,
         bd_spots = case_when(diff < -sd(eng_flickr) ~ "Dark",
                               diff > sd(eng_flickr) ~ "Bright",
                               TRUE ~ "Neutral"))

# potentially explanatory spatial layers
# np <- st_read("~/DATA/ADMINISTRATIVE/national_parks_england/National_Parks_England.shp",
#               quiet = TRUE) %>%
#   st_transform(st_crs(study_ext_sf))
# 
# aonb <- st_read("~/DATA/ADMINISTRATIVE/aonb_england/Areas_of_Outstanding_Natural_Beauty_England.shp",
#                 quiet = TRUE) %>%
#   st_transform(st_crs(study_ext_sf))

cities <- st_read("~/DATA/ADMINISTRATIVE/uk_cities/Major_Towns_and_Cities_December_2015_Boundaries.shp",
                  quiet = TRUE) %>%
  st_transform(st_crs(study_ext_sf)) %>%
  st_centroid %>%
  filter(tcity15nm %in% c("Birmingham", "Leeds", "London", "Manchester", "Newcastle upon Tyne"))

ggplot() + 
  geom_raster(data = bd_spots, 
              aes(x = x, y = y, fill = bd_spots)) + 
  #geom_sf(data = np, colour = "black", fill = NA) + 
  #geom_sf(data = aonb, colour = "black", fill = NA) + 
  geom_sf(data = cities, colour = "black", fill = NA) + 
  geom_sf(data = study_ext_sf, fill = NA) + 
  facet_wrap(~resolution) + 
  scale_fill_manual(values = c("orange", "blue", "grey")) + 
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

Cities marked out: London, Birmingham, Manchester, Leeds, Newcastle upon Tyne.  
```{r}
ggplot(bd_spots, aes(x = mod_fitted, y = eng_flickr)) + 
  geom_point(aes(colour = bd_spots)) + 
  scale_colour_manual(values = c("orange", "blue", "grey")) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~resolution, scales = "free")
```

Again problems with the 1km model.

Scale dependency in the % of bright/dark spots?

```{r}
bd_summary <- group_by(bd_spots, resolution) %>% 
  summarise(Bright = sum(bd_spots == "Bright"),
            Dark = sum(bd_spots == "Dark"),
            Total = n(),
            `Bright %` = (Bright/Total)*100,
            `Dark %` = (Dark/Total)*100) %>% 
  gather(key, value, -resolution, -Bright, -Dark, -Total)

ggplot(bd_summary, aes(x = resolution, y = value, colour = key, group = key)) + 
  geom_point() + 
  scale_colour_manual(values = c("orange", "blue", "grey")) + 
  stat_summary(fun.y=sum, geom="line") + 
  ylab("% of total area")
```

## Hot & cold spots

These are the areas of maximum and minimum service provision. I'm defining hot/cold spots as those cells in the top/bottom 5%. 

I'm not 100% convinced by this because of the skewed distribution. 

```{r}
hc_spots <- df %>% 
  group_by(resolution) %>% 
  mutate(hc_spots = case_when(eng_flickr < quantile(eng_flickr, 0.1) ~ "Cold",
                               eng_flickr > quantile(eng_flickr, 0.9) ~ "Hot",
                               TRUE ~ "Neutral"))
```

```{r}
ggplot() + 
  geom_raster(data = hc_spots, 
              aes(x = x, y = y, fill = hc_spots)) + 
  #geom_sf(data = np, colour = "black", fill = NA) + 
  #geom_sf(data = aonb, colour = "black", fill = NA) + 
  geom_sf(data = cities, colour = "black", fill = NA) + 
  geom_sf(data = study_ext_sf, fill = NA) + 
  facet_wrap(~resolution) + 
  scale_fill_manual(values = c("blue", "orange", "grey")) + 
  theme(axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

Problem in that coastal areas come out as cold spots, but that's an area thing. Might need to remove cells which are not 100% on land. 

Scale dependency in the % of hot/cold spots?

```{r}
hc_summary <- group_by(hc_spots, resolution) %>% 
  summarise(Hot = sum(hc_spots == "Hot"),
            Cold = sum(hc_spots == "Cold"),
            Total = n(),
            `Hot %` = (Hot/Total)*100,
            `Cold %` = (Cold/Total)*100) %>% 
  gather(key, value, -resolution, -Hot, -Cold, -Total)

ggplot(hc_summary, aes(x = resolution, y = value, colour = key, group = key)) + 
  geom_point() + 
  scale_colour_manual(values = c("blue", "orange", "grey")) + 
  stat_summary(fun.y=sum, geom="line") + 
  ylab("% of total area")
```

## References

## Session Info

```{r session_info}
session <- devtools::session_info()
session[[1]]
session[[2]] %>% kable
```